---
title: "Deductible versus Coinsurance under an ADAS Loss Profile"
subtitle: "Track 4 project specification, MFI Refresher Group Presentation"
author:
  - name: "Hoyong Jung"
    email: hoyong.jung@mail.utoronto.ca
  - name: "Daren Wang"
    email: daren.wang@mail.utoronto.ca
  - name: "Yufei (Faye) Wang"
    email: fayeee.wang@mail.utoronto.ca
  - name: "Qiwei Ye"
    email: qiwei.ye@mail.utoronto.ca
  - name: "Xiaoxiao Fan"
    email: xiaoxiao.fan@mail.utoronto.ca
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    embed-resources: true
    theme: cosmo
    code-fold: show
    df-print: kable
execute:
  eval: false
  echo: true
  warning: false
---

## Purpose of this document

This is a working specification, not a report. It fixes the decisions that everyone
else's work depends on, so that five people can work in parallel without waiting on
each other.

Anything marked **fixed** should be treated as settled unless someone raises a
substantive objection. Anything marked **open** is genuinely undecided and needs an
owner.

**Presentation: Monday 14 September 2026, 18:00 to 21:00, UY 9016.** This is the first
STA 2550 session on the registrar timetable. An earlier email mentioned 9 September,
but that falls on a Wednesday and STA 2550 meets on Mondays, so 14 September is the
date to plan around.

A note on numbering. This document numbers its own sections, and the presentation has
five sections of its own. They are different things. Whenever this document refers to
one of the five presentation sections it says so explicitly, for example "presentation
section 4".

### What changed in this revision, 7 September

Nobody needs to reread the whole document. These are the substantive changes.

**From the validation run (`results/data_processing.md`) and the claim-table checks
(`results/claim_table_checks.md`):**

- **13.1% of claims are non-positive.** 597 negative and 2 zero out of 4,568. Cleaned
  claim count is 3,969 and $\lambda$ falls about 13%. Limitation 8 rewritten.
- **The negatives are explained.** Per the CASdatasets documentation they are claims
  where our insured's liability was not engaged, recovered through legal recourse.
  Excluding them is a scope decision, not a cleaning convenience, and $\lambda$ is now
  an at-fault frequency. Two earlier hypotheses failed and are kept on the record.
- **The claim table is one row per accident, but the earlier reason was wrong.** The
  date-uniqueness check had been run on the *policy* table. The claim table's date
  combination is **not** unique. The real evidence is `ClaimNb`: 6 disagreements out of
  4,244 groups.
- **There is no accident date in the data.** The claim table's dates are the policy
  period copied across, matching at 100%.
- **The supplied `ClaimNb` column must not be used** after cleaning. It reconciles to
  the uncleaned 4,568. Rebuild counts from the cleaned claim table.
- **The exposure mismatch is unresolved, and the vehicle-years guess was wrong.** The
  documentation defines `Exposure` as period length over 365, so the pipeline check is
  right and the data does not match its own documentation. We keep the supplied column
  regardless.
- **`SumInsured` is a banded category, not a number**, so $L$ takes a band edge rather
  than a value read off the field.
- **The train-only categorical levels are confirmed harmless**: train is the superset,
  and we use no covariates.

**Decisions and corrections:**

- **Install path corrected**; `xts`/`zoo` were missing from the prerequisites.
- **The data is three years (2011-2013), not one.** Limitation 6 rewritten.
- **A policy row is a covered vehicle over a period, not a policy.** Do not merge or
  split. The claim table is already one row per accident; do not aggregate it.
- **Severity multiplier changed from 1.10 to 1.15.** The PDL severity figure exists in
  the same HLDI source, so both multipliers now come from the coverage that matches
  our data.
- **Portfolio size fixed at 1,000 vehicles.**
- **Per-claim limit $L$ now has a source and a robustness test**, after Xiaoxiao's
  review.
- **Simulation code corrected, and a Python translation note added.** Both negative
  binomial parameters scale with portfolio size. In `scipy.stats.nbinom` terms, only
  $n$ scales and $p$ does not.
- **Three slides each, not four.** Four each did not fit the 20-slide cap.
- **Assumptions split from limitations** between presentation sections 2 and 5.

**Numbers freeze target: Wednesday 9 September**, subject to Qiwei's timing. Full
run-through with timing and practice questions Saturday 12 September.

---

## Central question

> **Fewer accidents but more expensive repairs. For an insurer facing that loss
> profile, is a higher deductible or a lower deductible with coinsurance the better
> contract design?**

Status: **fixed**.

Two things are worth saying about the shape of this question.

First, it is a *decision* question, not a survey question. The brief is explicit that
a presentation should investigate a decision or problem rather than survey a topic.
We compare exactly two contract designs and defend a choice between them.

Second, the automation angle is the *motivation*, not the subject. We are not building
a model of self-driving cars. We are taking a real, documented shift in the loss
profile of vehicles equipped with driver assistance systems, and asking what it
implies for contract design. That distinction keeps the scope inside what we can
defend.

### Modules connected

| Module | Role |
|---|---|
| P&C insurance | Primary. Compound frequency-severity model, loss transformation, tail measures, truncation and censoring. |
| Programming fundamentals | Implementation. Reproducible pipeline, functions, Monte Carlo simulation, this document itself. |

The brief allows programming to serve as the implementation layer rather than a
separate theoretical subject. It still has to be substantive, which is why the
pipeline is written as functions in a versioned repository rather than as a notebook
that only one person can run.

---

## Dataset decision

Status: **fixed**.

Qiwei prepared a six-dataset public inventory, scoring each candidate against eighteen
variable categories required or useful for this simulation. That inventory drives the
decision below, and it should be credited on the reference slide.

| Dataset | Coverage | Size | Score |
|---|---|---|---|
| **CAS Pricing Game 2016 (`pricingame`)** | French private motor, material damage | 87,228 policies + 4,568 claims | **12/18** |
| Spanish actual motor portfolio 2015-2018 | Motor, aggregated by policy period | 105,555 policy-period rows | 11/18 |
| Spanish coverage-specific portfolio 2022-2024 | Motor, coverage level | 354,140 policy-years | 11/18 |
| Brazilian `brvehins1` | Vehicle insurance, aggregated | 1,965,355 records | 7/18 |
| freMTPL2 | Motor third-party liability | 677,991 policies + 26,444 claims | 6/18 |
| Guillen-Bolance-Frees-Valdez five-year panel | Motor and home, annual aggregates | 122,935 customer-years | 6/18 |

### Selection: `pg16trainpol` and `pg16trainclaim`

We use the 2016 Pricing Game tables from the `CASdatasets` package. Three reasons.

**Highest coverage of the variables we need**, at 12 of 18. It is the only candidate
that carries deductible and sum-insured information alongside claim-level amounts,
which lets us check that our chosen contract terms sit in a plausible range rather
than being invented.

**Material damage only, bodily injury excluded.** This matters more than the score.
Our question is about repair costs. A severity distribution that mixes vehicle damage
with bodily injury is a mixture of two very different processes, and fitting a single
parametric family to it would be hard to defend. Excluding injury gives us a cleaner
repair-cost distribution.

**Claim-level amounts, not aggregates.** Four of the six candidates report losses
aggregated over a policy period or a year. We need individual claim amounts to fit a
severity distribution and to apply a per-claim deductible.

### Why not freMTPL2

It was the initial suggestion and it is wrong for this question. It is the best-known
teaching dataset in this area, which is exactly why it was suggested, but familiarity
is not fitness. Third-party liability severities include bodily injury, and the public
files carry no deductible or limit information at all. Qiwei's inventory caught this;
we are following the evidence rather than the default.

### Loading

The package is not on CRAN and the path in an earlier version of this document was
stale. `xts` and `zoo` must be installed first or the load fails with an unhelpful
error.

```{r}
#| label: load-data
install.packages(c("xts", "zoo"))

# Three mirrors, try in order. Large download, allow a few minutes.
install.packages("CASdatasets",
                 repos = "https://dutangc.perso.math.cnrs.fr/RRepository/pub/",
                 type = "source")
# or repos = "http://dutangc.free.fr/pub/RRepos/pub/"
# or repos = "https://cas.uqam.ca/pub/"
# last resort: devtools::install_github("dutangc/CASdatasets")

library(CASdatasets)
data(pricingame)

pol   <- pg16trainpol    # 87,228 policy-period rows
claim <- pg16trainclaim  # 4,568 claims, individual amounts

str(pol)
str(claim)
```

Two practical notes for whoever handles the data. Most categorical levels are
anonymised for confidentiality, so they carry no interpretable meaning and should not
be used to tell a story about driver or vehicle type. And the join between the two
tables is on the policy identifier, so policies with no claim will not appear in the
claim table at all.

### What a row actually is

Status: **fixed**, settled against the data 6-7 September and confirmed by the
validation run in `results/data_processing.md`.

**The policy table is not one row per policy.** A row is one covered vehicle over one
coverage period. A single `PolicyID` can appear several times: once per vehicle
(`LicNb`) and once per period when cover is renewed or amended. Example from the data,
`PolicyID` 3933 occupies four rows covering two vehicles across three years. The
validation confirms `(PolicyID, LicNb, Year, BeginDate, EndDate)` is unique with zero
duplicates, so that tuple is the row key.

Leave those rows as they are. Do not merge them into one row per policy and do not
split them further. The frequency model uses exposure as an offset, so the unit that
matters is vehicle time on risk, not the count of policy numbers. Each row is one
observation carrying its own exposure, and claims attach to the row they fall in.
This is the standard vehicle-year construction.

**The training data covers three years, 2011 to 2013.** A fourth year, 2014, is held
out as `pg16test` (32,772 rows), which we do not use. An earlier version of this
document assumed a single year, and a later one said four; both were wrong. The
variable inventory in `results/data_processing.md` shows three distinct `Year` values
in training and one in test. We pool all three years. The exposure offset already puts
everything on an annual basis, so pooling still yields an annual $\lambda$ and gives us
three times the data.

**The claim table is one row per accident.** Verified on 7 September and recorded in
`results/claim_table_checks.md`.

An earlier version of this document justified this by claiming
`(PolicyID, LicNb, date)` is unique in the claim table. **That was wrong.** The
uniqueness check it leaned on had been run on the *policy* table, and
`(PolicyID, LicNb, Year, BeginDate, EndDate)` is in fact **not** unique in the claim
table: 604 rows fall into 286 repeated groups.

The correct justification is the insurer's own count. Grouping claim rows by
`(PolicyID, LicNb, Year)` and comparing against `ClaimNb` on the matching policy
sub-period, **6 groups out of 4,244 disagree and all by exactly one**. Totals reconcile
at 4,568 = 4,568. If the repeated rows were split payments on a single accident,
`ClaimNb` would sit *below* the row count by roughly the 318 surplus rows. It does not.
The insurer is counting each claim row as a separate accident, so we do too, and we do
not aggregate.

`results/claim_table_checks.md` contains a line reading the duplicates as payment rows.
Disregard it; it contradicts the reconciliation in the same document.

**There is no accident date in the data.** `BeginDate` and `EndDate` in the claim table
are the policy coverage period copied across from the policy row, matching at 100.00%
on all 4,568 rows. Two accidents falling in the same coverage period are therefore
indistinguishable from each other, which is why the date could never have settled this
question. It does not matter for us: an intercept-only frequency model needs the count
per exposure period, not the day each accident happened. Say this if asked what is
missing from the data.

### The exposure mismatch

Status: **unresolved, and we use the supplied column anyway.** This was Qiwei's
original point 2.

The validation flags 82,435 of 87,228 rows where the supplied `Exposure` differs from
`(EndDate - BeginDate) / 365`, with 18,052 rows off by more than a rounding tolerance
and a maximum gap of 0.166.

The CASdatasets documentation defines `Exposure` as exactly that quantity: the
difference between `EndDate` and `BeginDate` divided by 365. **So the pipeline check
implements the documented formula correctly, and the data still does not satisfy it.**
The discrepancy is in the file, not in our code.

An earlier version of this document guessed that `Exposure` was vehicle-years,
`VehiclNb` times the period length, on the strength of three examples whose ratios were
2.01, 2.05 and 1.01, and the fact that `VehiclNb` only takes values 1 and 2. The
documentation rules that out as the *definition*. It may still describe what the
numbers happen to be, and one line would tell us:

```python
(pol["Exposure"] / (pol["VehiclNb"] * pol["ExposureFromDates"])).describe()
```

If that clusters at 1.0, then the supplied values were computed on a vehicle basis
despite the documented formula, and we can say so. If it does not, we simply record
that the documented formula does not reproduce the column.

**Either way we keep the supplied `Exposure`.** It is the insurer's own measure of time
on risk, and a derived quantity that contradicts the data's own documentation is not a
better substitute. Note that this rescales $\lambda$ by a few percent at most, while
our scenario shifts frequency by 40%, so it cannot change the conclusion.

### Cleaning the claim amounts

Status: **fixed in rule, one diagnostic outstanding**.

The validation found **597 negative and 2 zero `ClaimCharge` values out of 4,568**.
That is **13.1% of the claim table**, which is far more than the handful this document
previously assumed. It is large enough to be a headline limitation rather than a
footnote. See limitation 8.

Lognormal and gamma need strictly positive support, so the 599 rows come out of the
severity fit. Two consequences follow and neither is optional:

1. **They come out of the claim counts as well.** After cleaning the count is 3,969,
   not 4,568, so $\lambda$ falls by about 13%. The `ClaimNb` column in the policy table
   currently reconciles to the uncleaned 4,568 and must be rebuilt by aggregating the
   cleaned claim table back onto policy rows. Using the supplied `ClaimNb` alongside a
   cleaned severity fit would mean the simulation draws 4,568 claims and assigns every
   one a positive cost, which inflates the aggregate.
2. **The drop is not random.** We are removing accidents on which the insurer ended up
   net positive, so the surviving severity distribution is conditional on the insurer
   having net paid. That pushes mean severity up while frequency goes down.

**The negatives are claims where our insured was not at fault.** The CASdatasets
documentation states, for the 2017 Pricing Game's `claim_amount` field, that negative
values arise from claims where the driver's liability is not engaged, so there is legal
recourse. The insurer spends on handling the claim, then recovers from the at-fault
party's insurer and can end up net positive on that accident.

**This is an interpretation carried across, and we say so.** The note appears under the
PG17 field, not under our PG16 `ClaimCharge`, whose entry does not explain the sign.
What the PG16 documentation *does* record is the French IDA convention: the insurer may
refund its own insured directly (`DirectComp = TRUE`) and afterwards sue the third-party
insurer to recover. That is the same mechanism, in the same market, in the same data
family. Adopting the PG17 explanation for PG16 is reasonable and should be presented as
a reading rather than a documented fact.

Two earlier guesses failed and are worth keeping on the record, because "we tested two
explanations and neither held, then found the documentation" is a stronger story than
arriving at the answer directly:

- *Direct-compensation recoveries identified through `DirectComp`.* 570 of the 597
  negatives sit at `DirectComp = False`, which looked like a refutation. It is not.
  `DirectComp = TRUE` marks direct refund with later recovery, and the negative rate
  there is 21% against 13% overall, pointing the right way on 126 rows. Recourse can
  also proceed outside the IDA route. The cross-tab was never decisive either way.
- *Reserve movements on unsettled claims, using `SettlYear = 0` as the marker.*
  Abandoned. The documentation does not explain what `SettlYear = 0` means, so the
  settled/unsettled split it assumed is unfounded.

**Dropping them is now correct by definition, not by convenience.** These are not
losses our insured caused, and our question is what our insured's accidents cost. The
cleaned count of 3,969 is at-fault accidents, which is also the right basis for the ADAS
scenario, since that scenario is about our vehicles causing fewer crashes.

Record and report: rows in, rows out, how many zero, how many negative.

---

## The ground-up gap

Status: **fixed as a limitation**. This is the most important assumption in the project
and it belongs on a slide.

Qiwei's matrix has a column labelled GU: verified ground-up repair cost, before any
deductible, coinsurance or limit is applied.

**That column is zero for all six candidates.** No public dataset in the inventory
publishes a verified ground-up loss.

This is not a reason to keep searching. It is a structural property of insurance data.
An insurer only observes losses that were reported to it, and a policyholder has no
reason to report a loss smaller than their deductible. The data are therefore already
**left-truncated** at an unknown threshold, and any distribution fitted to them is a
conditional distribution dressed up as an unconditional one.

The consequence for us is specific and quantifiable:

- Fitting a severity distribution to observed claims and treating it as ground-up
  **understates the true frequency** of small losses.
- That in turn **overstates the cost** of the low-deductible design, because the small
  losses that a low deductible would newly pick up are missing from the data.
- The bias therefore runs in a known direction, and it favours the high-deductible
  design.

Stating the direction of the bias is much stronger than stating that a bias exists. It
also answers one of the questions every student is told to be ready for: what
information is missing, censored, or aggregated.

The Track 4 brief lists discussion of censoring and truncation as required evidence.
So the gap Qiwei found is not an obstacle to the project. It is one of the
deliverables.

---

## Model specification

Status: **fixed**, subject to the diagnostics below coming out as expected.

We use a collective risk model. For a portfolio over one year,

$$S = \sum_{i=1}^{N} X_i$$

where $N$ is the number of claims, $X_i$ are individual ground-up loss amounts, and
$N$ and $X_i$ are assumed independent.

### Frequency

**Negative binomial**, with Poisson fitted alongside as a diagnostic.

The rationale is overdispersion. A Poisson model forces the variance to equal the mean,
and motor claim counts almost always show a variance above the mean because
policyholders are not identical. The negative binomial arises naturally as a Poisson
mixture over unobserved heterogeneity, which is the right story here given that the
risk characteristics in this dataset are anonymised and cannot fully explain that
heterogeneity.

Fit both, report the dispersion statistic, and keep Poisson in the appendix. If the
data turn out not to be overdispersed, we use Poisson and say so.

Exposure enters as an offset, since rows are observed for varying fractions of a year.
With three years pooled, the offset is what converts everything back to an annual basis,
so the fitted $\lambda$ is an annual claim rate per vehicle-year regardless of how many
calendar years went in.

One consequence of pooling worth stating if asked: the same policy contributes several
rows, so the rows are not strictly independent. The negative binomial absorbs some of
this, since repeat observations of one policy are exactly the unobserved heterogeneity
the mixture is there to represent, but it is not a full panel treatment. See limitation
6.

### Severity

**Lognormal and gamma, choose by fit.**

Both are standard for positive, right-skewed loss amounts. Lognormal has the heavier
right tail of the two, which matters for a question about tail risk. Select by AIC
supported by a QQ plot, and show the QQ plot rather than only the AIC number, because
a number is not interpretable and a plot is.

We do **not** use a Pareto-type severity. It would be defensible, but with 4,568 claims
the tail index would be poorly estimated, and the brief is explicit that a simple
analysis that is correctly understood beats a complex one that cannot be defended.

### What we are not doing

No GLM with covariates. No machine learning. No traffic simulation. No correlated or
common-shock structure between claims. Each of these was considered and dropped for
scope; if asked, the answer is that the brief limits us to one main dataset and two
principal methods, and that the question does not require them.

### Implementation note: the pipeline is Python

Status: **fixed**. The code blocks in this document are written in R because that is
how the modelling decisions are usually expressed, but the actual pipeline is Python
(`src/track_4/`, run with `uv run src/run_pipeline.py`, outputs to Parquet). The
decisions are language-independent. One translation is not, and it is the one that can
go wrong silently.

**`rnbinom(mu, size)` has no direct Python equivalent.** Both
`numpy.random.negative_binomial` and `scipy.stats.nbinom` use $(n, p)$. Convert with

$$n = \theta, \qquad p = \frac{\theta}{\theta + \mu}$$

where $\mu$ and $\theta$ are the fitted per-vehicle mean and dispersion. When scaling
to a fleet of 1,000 vehicles, **only $n$ scales**. The probability $p$ is a ratio and
both parameters scale together, so it is unchanged:

```python
from scipy.stats import nbinom
p = theta / (theta + mu)
N = nbinom.rvs(n = n_veh * theta, p = p, size = n_sims, random_state = seed)
```

Scaling $\mu$ while leaving $n$ at the per-vehicle value gives variance
$n\mu + (n\mu)^2/\theta$, which is wrong by orders of magnitude and inflates every tail
measure without producing any visible error. This is the same bug described in the
contract designs section, in a different language.

For the fits themselves, `statsmodels.discrete.discrete_model.NegativeBinomial` with an
`exposure` argument replaces `glm.nb`, and `scipy.stats.lognorm.fit` /
`scipy.stats.gamma.fit` replace `fitdist`. Compare by AIC as planned.

```{r}
#| label: fit
# Reference implementation. See the note above; the pipeline is Python.
library(fitdistrplus)
library(MASS)

# One cleaned claim table drives BOTH the counts and the severity fit.
claim_clean <- subset(claim, ClaimCharge > 0)   # drops 599 of 4,568 rows

n_zero <- sum(claim$ClaimCharge == 0)   # 2
n_neg  <- sum(claim$ClaimCharge <  0)   # 597, see limitation 8

# Counts rebuilt from the cleaned claims. Do NOT use the supplied ClaimNb column,
# which reconciles to the uncleaned 4,568.
counts <- aggregate(list(ClaimNbClean = claim_clean$ClaimCharge),
                    by  = list(PolicyID = claim_clean$PolicyID,
                               LicNb    = claim_clean$LicNb),
                    FUN = length)
pol <- merge(pol, counts, by = c("PolicyID", "LicNb"), all.x = TRUE)
pol$ClaimNbClean[is.na(pol$ClaimNbClean)] <- 0

# Frequency. Exposure is vehicle-years and is used as supplied.
fit_pois <- glm(ClaimNbClean ~ 1, offset = log(Exposure), family = poisson, data = pol)
fit_nb   <- glm.nb(ClaimNbClean ~ 1 + offset(log(Exposure)),                 data = pol)

# Severity
fit_ln <- fitdist(claim_clean$ClaimCharge, "lnorm")
fit_ga <- fitdist(claim_clean$ClaimCharge, "gamma")

gofstat(list(fit_ln, fit_ga))
```

The claim table has no `LicNb`-level period, so the join above attaches a claim to a
vehicle rather than to a specific coverage period. If a vehicle has several periods in
the pool, decide whether to join on date range as well; with an intercept-only model
this affects nothing, since the totals are the same either way, but say which you did.

---

## Scenario multipliers

Status: **fixed for the baseline, sensitivity grid open**.

We do not invent the frequency and severity shift. It comes from published insurance
loss data, which is what makes this project defensible rather than a parameter
guessing exercise.

The Highway Loss Data Institute compared 2017-22 model year vehicles with and without
a bundle of driver assistance features, holding the vehicle model otherwise constant.
The findings:

| Quantity | Effect for equipped vehicles |
|---|---|
| Collision claim severity | about **10% higher** |
| Collision claim frequency | about **10% lower** |
| **Property damage liability claim severity** | about **15% higher** |
| **Property damage liability claim frequency** | about **40% lower** |
| Overall collision losses | about **5% lower** |

The mechanism is that sensors are expensive to replace and usually require calibration
afterwards, but equipped vehicles are in fewer crashes to begin with.

### Baseline scenario

Our data are third-party material damage, so the property damage liability figures are
the closer match. **Both multipliers are taken from PDL coverage**, so they come from
the same coverage in the same study rather than being mixed across coverages.

- Frequency multiplier $\phi_N = 0.60$
- Severity multiplier $\phi_X = 1.15$

Applied as: $\lambda_{AV} = \phi_N \lambda$, and severity scaled by $\phi_X$ on the
mean while holding the coefficient of variation constant.

An earlier version of this document used $\phi_X = 1.10$, the collision figure, and
then apologised for mixing coverages. The PDL severity figure exists in the same
source; using it removes the mismatch.

### Sensitivity grid

Run the full model over $\phi_N \in [0.5, 1.0]$ and $\phi_X \in [1.0, 1.3]$ and present
a contour of the result. This is the single most valuable slide in the deck, because it
converts our biggest weakness, dependence on assumed parameters, into the subject of
the analysis.

### What the severity uplift actually measures

Status: **fixed**. Owned by presentation section 5.

Now that both multipliers come from PDL, the coverage mismatch is gone. Two things
remain worth saying, and the second is more interesting than the one it replaces.

**Country.** The multipliers are US experience; the data are French. Repair cost
structures and liability rules differ. We keep this in the limitations register.

**Mechanism.** HLDI notes that the PDL severity gap is larger than sensor replacement
costs alone can explain, and attributes part of it to crash avoidance removing slower,
low-cost crashes from the claim mix, which raises the average of what remains. That
matters for how we apply the uplift. We scale the mean and hold the coefficient of
variation constant, which implicitly treats the whole uplift as every loss inflating.
If part of it is a mix shift, the left tail is being removed instead, which raises the
mean **and lowers** the CV.

The direction is worth stating because it is useful to us. A genuine mix shift removes
exactly the small losses that a low deductible would newly pick up, so it works in
favour of Design B. That runs opposite to the ground-up truncation bias in the section
above, which favours Design A. The two partially offset, and being able to say so is a
stronger answer than naming either bias alone.

---

## Contract designs

Status: **fixed in structure, parameters open pending calibration**.

Insurer payment on a ground-up loss $X$:

$$Y = \min\big(\alpha (X - d)_+,\; L\big)$$

where $d$ is the deductible, $\alpha$ the insurer's share above the deductible, and $L$
the per-claim limit.

| | Design A: high deductible | Design B: coinsurance |
|---|---|---|
| Deductible $d$ | 1,000 | 250 |
| Insurer share $\alpha$ | 1.00 | 0.80 |
| Per-claim limit $L$ | from a `SumInsured` band edge | same value |

### Portfolio size

Status: **fixed**. Simulate a fleet of **1,000 vehicles**, not the full 87,228 rows.

This was previously unspecified and it is not a detail. At full portfolio size the
aggregate diversifies, $\mathrm{TVaR}_{99}/E[S]$ sits close to 1, and the difference
between the two designs disappears into rounding. A thousand-vehicle commercial fleet
account keeps the tail a tail, and it matches the framing of the central question.
State the number on the slide.

### The per-claim limit

Status: **fixed**. Raised by Xiaoxiao on 6 September.

$L$ previously read "same" in both columns, which is not a value.

**`SumInsured` is a banded category, not a number.** The validation lists levels such
as `<=7.6 Keur`, so it cannot supply a precise $L$. It can still anchor one: take the
upper edge of a high band as the limit and say that is where it came from, which is
better than inventing a round figure. Fall back to the 99.5th percentile of the cleaned
claim amounts if the bands turn out unusable, and say which you used. `Deduc` is
banded the same way and is worth checking for a different reason: it tells us whether
$d = 1{,}000$ and $d = 250$ sit inside the range of deductibles the portfolio actually
sold, which is the check this dataset was chosen to make possible.

$L$ caps the **insurer payment**, not the ground-up loss, which matters because
$\alpha \neq 1$ in Design B.

**A common $L$ is not a common effect.** Design A reaches the cap at a ground-up loss
of $L + 1{,}000$; Design B reaches it at $1.25L + 250$. A hits the ceiling at a
*smaller* loss, so A's tail is truncated first and A can look better on TVaR for a
reason that has nothing to do with risk sharing. This runs against the structural
expectation described below, so it has to be tested rather than assumed away.

Two things settle it:

1. Report the share of claims hitting the limit under each design. This is already in
   the required outputs but was not being read as a diagnostic.
2. **Run the whole comparison a second time with no limit at all.** If the ranking is
   unchanged with and without, one line on the slide closes the question. If it
   changes, that is the finding and we present it as such.

### What we expect to find, and why we say so in advance

A deductible subtracts a fixed amount, so its relief becomes negligible as losses grow.
Coinsurance is proportional, so it keeps clipping the same share however far into the
tail you go. At equal expected cost, Design B should therefore show the lower
$\mathrm{TVaR}_{99}$ for any right-skewed severity. Only a binding limit changes that,
which is what the test above is for.

This is a mechanism we can explain at a whiteboard, not a discovery. Present it that
way. It also means the sensitivity contour may not flip anywhere in the grid: check
that on a coarse 4x4 run before building the slide, and if there is no flip, plot the
ratio $\mathrm{TVaR}_{99}(A)/\mathrm{TVaR}_{99}(B)$, which varies, instead of a binary
preference, which would be one solid colour.

### The calibration that makes the comparison fair

**Calibrate the two designs to approximately equal expected insurer cost at baseline,
then compare their tail behaviour and their effect on the policyholder.**

Without this step the comparison is trivial and the presentation is weak: of course a
higher deductible costs the insurer less. Holding expected cost equal turns the
question into something that actually has an answer worth defending, which is how the
two designs distribute risk rather than how much they cost on average.

Tune $d$ in Design A, or $\alpha$ in Design B, until expected aggregate cost matches
within a small tolerance. Report the calibrated values and the tolerance achieved.

**Both negative binomial parameters scale with portfolio size.** For $n$ independent
vehicles each $\mathrm{NB}(\mu, \theta)$, the portfolio total is
$\mathrm{NB}(n\mu,\; n\theta)$. Putting a portfolio-scale mean against a per-vehicle
dispersion gives variance $n\mu + (n\mu)^2/\theta$, which is wrong by orders of
magnitude and would silently inflate every tail measure. An earlier version of this
code block had exactly that error.

```{r}
#| label: payment
insurer_payment <- function(x, d, alpha, limit) {
  pmin(alpha * pmax(x - d, 0), limit)
}

# rsev is a closure, e.g. function(k) rlnorm(k, ml, sdl) or rgamma(k, shape, rate),
# so switching the fitted severity family is a one-line change.
simulate_aggregate <- function(n_sims, n_veh, mu_veh, theta_veh,
                               rsev, d, alpha, limit = Inf) {
  N   <- rnbinom(n_sims, mu = n_veh * mu_veh, size = n_veh * theta_veh)
  tot <- sum(N)
  agg <- numeric(n_sims)
  if (tot > 0) {
    y   <- insurer_payment(rsev(tot), d, alpha, limit)
    idx <- rep.int(seq_len(n_sims), N)
    s   <- tapply(y, idx, sum)
    agg[as.integer(names(s))] <- s     # sims with zero claims stay at 0
  }
  agg
}

tvar <- function(s, p = 0.99) mean(s[s >= quantile(s, p)])

# Monte Carlo standard error on the tail measure
tvar_se <- function(s, p = 0.99) {
  tail <- s[s >= quantile(s, p)]
  sd(tail) / sqrt(length(tail))
}
```

Set a seed and run **500,000 simulations**, not 100,000. At 100k only about 1,000
draws sit above $\mathrm{VaR}_{99}$, so the standard error on $\mathrm{TVaR}_{99}$ is
not negligible. Vectorised, 500k costs seconds. Report the standard error alongside
every tail figure; a tail estimate quoted without its own uncertainty invites an
obvious question.

---

## Required outputs

Status: **fixed**. Every item below appears either on a slide or in the appendix.

**Aggregate loss distribution**, both designs, both baseline and shifted scenarios:

- Expected aggregate insurer cost, $E[S]$
- Standard deviation
- VaR at 95% and 99%
- **TVaR at 99%** (the headline tail measure)

**Policyholder side:**

- Expected retained loss per policyholder, $E[X - Y]$
- Retention as a share of ground-up loss

**Truncation and censoring diagnostics** (required by the Track 4 brief):

- Share of claims falling entirely below the deductible under each design
- Share of claims hitting the limit, **reported separately for each design**
- Rows dropped in cleaning: how many zero, how many negative
- How each design changes what the insurer would observe in future data

**Sensitivity:**

- Contour over the $(\phi_N, \phi_X)$ grid. Plot the ratio
  $\mathrm{TVaR}_{99}(A)/\mathrm{TVaR}_{99}(B)$ rather than a binary preference unless
  the coarse check shows the preference actually flips somewhere in the grid.
- **Unlimited run.** The full comparison repeated with $L = \infty$, to show the
  ranking is not an artefact of the common limit.

Keep every figure legible from the back of a room. Two well-labelled charts beat six
crowded ones, and the deck is capped at 20 slides including cover and figures.

---

## Who owns what

Five sections in the brief, five of us. Each person owns one presentation section and
presents it, two to four minutes each.

**Three slides each, maximum.** Four each would be 20 plus a cover plus a references
slide, which is 22 against a cap of 20 including the cover. Three each gives 15, plus
cover and references, for 17, and leaves room for the sensitivity figure to take a full
slide.

| Presentation section | Owner | Needs |
|---|---|---|
| 1. Problem and decision | Yufei (Faye) Wang | nothing, can start now |
| 2. Key concepts | Hoyong Jung | nothing, can start now |
| 3. Analysis and evidence | Qiwei Ye | the dataset and model spec above |
| 4. Comparison and judgment | Daren Wang | outputs from presentation section 3 |
| 5. Conclusion and limitations | Xiaoxiao Fan | the ground-up gap and limitations register |

### A note for presentation sections 4 and 5

Both sit downstream of the simulation, which creates a scheduling risk: if you wait for
section 3 to finish, you compress into the final days.

Neither of you actually needs to wait.

**Section 4** can be written against the *structure* of the result rather than the
numbers. The comparison logic, which measures matter and why, and what would make each
design preferable is all fixed by the contract designs section above. Draft it with
placeholder numbers.

**Section 5** can be drafted almost entirely from this document. The limitations
register below is the raw material, and it will not change when the simulation
finishes.

### Splitting assumptions from limitations

Status: **fixed**. Proposed by Xiaoxiao on 6 September, adopted with one change.

The limitations register mixes two different things: assumptions any actuarial model
would make, and assumptions whose violation moves our headline number in a direction we
can name. Presentation section 2 takes the first kind, stated once and briefly, so that
section 5 can go deeper on the second kind.

The dividing test is whether a reader could say **which way** the assumption pushes our
result.

**To presentation section 2, as an assumptions slide:** no policyholder behaviour
(a scope boundary, not a signed bias); multi-year pooling with no inflation adjustment;
independence stated plainly as part of defining the collective risk model.

**Staying with presentation section 5, as signed biases:** the ground-up gap; the
country mismatch; driver assistance is not full automation; independence *again*, but
as a directional bias rather than a definition, because correlated sensor or software
failure makes the real aggregate tail fatter than what we report and therefore makes
our TVaR optimistic; and expenses and recoveries, pending the count of negative claim
rows.

Section 2 defines the assumption; section 5 says what it costs us. Hoyong writes the
section 2 version so it sets section 5 up rather than repeating it.

### Review

Xiaoxiao has hands-on P&C experience and has agreed to review the modelling choices,
mainly the frequency and severity distributions and the tail measures, and to add
anything the limitations register is missing from a practitioner's point of view. This
sits alongside presentation section 5 rather than on top of it.

Her 6 September review produced the per-claim limit test now written into the contract
designs section, and the assumptions split above. Both are adopted.

### Where things live

Code and this specification live in the GitHub repository. Documents, slides, the
brief and any other shared files live in the Google Drive folder that Faye set up.
Nobody has to learn a new tool.

### One shared deck

**Owner: Faye.** Not a design exercise. The point is not aesthetics, which carry no
marks, but that five people building slides in five separate files produces a two-hour
merge on the Sunday night before. One Google Slides file, link shared, everyone adds
their own slides inside it. Whatever theme it opens with is fine. Needed before anyone
starts building.

Related: there are currently two Drive folders, Faye's and Qiwei's. Pick one and move
everything into it.

---

## Limitations register

Raw material for presentation section 5. Roughly in order of how badly each one would
hurt if a question landed on it.

1. **No ground-up losses.** No public dataset publishes verified ground-up repair cost.
   Observed claims are left-truncated at an unknown reporting threshold, so the fitted
   severity is conditional, small losses are undercounted, and the bias favours the
   high-deductible design. This is the ground-up gap described above.

2. **Coverage mismatch in the scenario.** Frequency and severity multipliers come from
   US collision and property damage liability experience; the data are French
   third-party material damage. Both the country and the coverage differ, as noted
   under scenario multipliers above.

3. **Driver assistance is not full automation.** HLDI measures vehicles with assistance
   bundles, not autonomous vehicles. We use it as the best available observed evidence
   for the direction and rough size of the shift, not as a measurement of automation.

4. **Independence assumed.** Claims are treated as independent. Vehicles sharing
   software or sensor hardware plausibly have correlated failure modes, which would
   fatten the aggregate tail beyond what we report. Our tail measures are therefore
   likely optimistic. We did not model this; it is the obvious next step.

5. **Anonymised categorical levels.** Risk characteristics in this dataset have no
   published meaning, so we cannot segment or interpret by risk class.

6. **Three years pooled, no inflation adjustment.** An earlier version of this document
   assumed a single year. The training data covers 2011 to 2013, with 2014 held out as
   test data that we do not use. We pool all three and let the exposure offset put frequency
   on an annual basis, but we apply no trend or repair cost index to severity. Repair
   costs have been rising faster than general inflation, so our fitted severity
   understates current levels, and the understatement grows with the age of the
   earliest year in the pool. Pooling also means one policy contributes several rows,
   so the observations are not strictly independent across time; the negative binomial
   absorbs part of this as unobserved heterogeneity but it is not a panel model.

7. **No policyholder behaviour.** We compute retention mechanically. In reality a
   higher deductible changes reporting behaviour and lapse rates, which would feed back
   into both frequency and portfolio composition.

8. **13% of claims are excluded as out of scope.** Claim handling expense is absent
   from the data outright. Separately, **597 negative and 2 zero `ClaimCharge` values
   out of 4,568, or 13.1%**, are excluded. Per the CASdatasets documentation these are
   claims where our insured's liability was not engaged, so the insurer recovered
   through legal recourse and came out net positive. They are not losses our insured
   caused, so excluding them is a scope decision rather than a data-cleaning
   convenience.

   Two caveats to state. The explanation appears under the 2017 Pricing Game's
   `claim_amount` field; our PG16 `ClaimCharge` entry does not explain the sign, though
   PG16 does document the IDA convention that produces the same effect. So this is a
   reading carried across, and we present it as one. And the cleaned count of 3,969 is
   **at-fault accidents**, so $\lambda$ is an at-fault frequency. That is the correct
   basis for a scenario about our vehicles causing fewer crashes, but the number is not
   comparable to a headline claim frequency that includes not-at-fault claims.

   Together with limitation 1 these are both properties of the dataset rather than of
   our method, which is why they share a slide.

---

## Questions we will be asked

The brief lists the questions every student should be ready to answer. Roughly six
minutes of questions across the group, and the checklist requires each of us to respond
without relying on slide text. Everyone should be able to answer these for their own
section.

- What is the central question, and why is it worth investigating?
- Why a negative binomial rather than a Poisson?
- Why lognormal rather than gamma, or rather than Pareto?
- Where did the frequency and severity multipliers come from?
- What is the most important assumption, and which way does it bias the result?
- Why did you hold expected cost equal between the two designs?
- What does TVaR at 99% mean in plain language?
- What is missing from your data that you would want before using this in practice?
  (Answer: ground-up losses; an accident date; and gross repair cost separated from
  recoveries, so the not-at-fault claims could be analysed rather than excluded.)

The second, third and sixth are the ones most likely to catch someone out, because they
are choices we made rather than facts we looked up.

---

## References

Sources to carry onto the final reference slide.

- Dutang, C. and Charpentier, A. *CASdatasets*: Insurance datasets. `pricingame`
  reference, French Institute of Actuaries Pricing Games.
  <https://dutangc.github.io/CASdatasets/reference/pricingame.html>
- Moore, M. (IIHS-HLDI), analysis published 11 August 2026, citing an HLDI comparison
  of 2017-22 model year vehicles with and without a driver assistance bundle. Source of
  both the collision figures and the property damage liability severity and frequency
  figures used in our baseline.
- Highway Loss Data Institute. *Compendium of HLDI collision avoidance research*,
  Bulletin Vol. 40, No. 4, April 2023.
- Ye, Q. *Public Insurance Dataset Candidates: Auto Deductible vs Coinsurance
  Simulation.* Six-dataset inventory with variable availability matrix.
- MFI Refresher Group Presentation Project, Student Project Brief, Track 4.
