---
title: "Deductible versus Coinsurance under an ADAS Loss Profile"
subtitle: "Track 4 project specification, MFI Refresher Group Presentation"
author:
  - name: "Hoyong Jung"
    email: hoyong.jung@mail.utoronto.ca
  - name: "Daren Wang"
    email: daren.wang@mail.utoronto.ca
  - name: "Yufei (Faye) Wang"
    email: fayeee.wang@mail.utoronto.ca
  - name: "Qiwei Ye"
    email: qiwei.ye@mail.utoronto.ca
  - name: "Xiaoxiao Fan"
    email: xiaoxiao.fan@mail.utoronto.ca
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    embed-resources: true
    theme: cosmo
    code-fold: show
    df-print: kable
execute:
  eval: false
  echo: true
  warning: false
---

## Purpose of this document

This is a working specification, not a report. It fixes the decisions that everyone
else's work depends on, so that five people can work in parallel without waiting on
each other.

Anything marked **fixed** should be treated as settled unless someone raises a
substantive objection. Anything marked **open** is genuinely undecided and needs an
owner.

Presentation date: **9 September 2026, 18:00 EST**, first STA 2550 class.

---

## Central question {#sec-question}

> **Fewer accidents but more expensive repairs. For an insurer facing that loss
> profile, is a higher deductible or a lower deductible with coinsurance the better
> contract design?**

Status: **fixed**.

Two things are worth saying about the shape of this question.

First, it is a *decision* question, not a survey question. The brief is explicit that
a presentation should investigate a decision or problem rather than survey a topic.
We compare exactly two contract designs and defend a choice between them.

Second, the automation angle is the *motivation*, not the subject. We are not
building a model of self-driving cars. We are taking a real, documented shift in the
loss profile of vehicles equipped with driver assistance systems, and asking what it
implies for contract design. That distinction keeps the scope inside what we can
defend.

### Modules connected

| Module | Role |
|---|---|
| P&C insurance | Primary. Compound frequency-severity model, loss transformation, tail measures, truncation and censoring. |
| Programming fundamentals | Implementation. Reproducible pipeline, functions, Monte Carlo simulation, this document itself. |

The brief allows programming to serve as the implementation layer rather than a
separate theoretical subject. It still has to be substantive, which is why the
pipeline is written as functions in a versioned repository rather than as a notebook
that only one person can run.

---

## Dataset decision {#sec-data}

Status: **fixed**.

Qiwei prepared a six-dataset public inventory on 21 August, scoring each candidate
against eighteen variable categories required or useful for this simulation. That
inventory drives the decision below, and it should be credited on the reference
slide.

| Dataset | Coverage | Size | Score |
|---|---|---|---|
| **CAS Pricing Game 2016 (`pricingame`)** | French private motor, material damage | 87,228 policies + 4,568 claims | **12/18** |
| Spanish actual motor portfolio 2015-2018 | Motor, aggregated by policy period | 105,555 policy-period rows | 11/18 |
| Spanish coverage-specific portfolio 2022-2024 | Motor, coverage level | 354,140 policy-years | 11/18 |
| Brazilian `brvehins1` | Vehicle insurance, aggregated | 1,965,355 records | 7/18 |
| freMTPL2 | Motor third-party liability | 677,991 policies + 26,444 claims | 6/18 |
| Guillen-Bolance-Frees-Valdez five-year panel | Motor and home, annual aggregates | 122,935 customer-years | 6/18 |

### Selection: `pg16trainpol` and `pg16trainclaim`

We use the 2016 Pricing Game tables from the `CASdatasets` package. Three reasons.

**Highest coverage of the variables we need**, at 12 of 18. It is the only candidate
that carries deductible and sum-insured information alongside claim-level amounts,
which lets us check that our chosen contract terms sit in a plausible range rather
than being invented.

**Material damage only, bodily injury excluded.** This matters more than the score.
Our question is about repair costs. A severity distribution that mixes vehicle damage
with bodily injury is a mixture of two very different processes, and fitting a single
parametric family to it would be hard to defend. Excluding injury gives us a cleaner
repair-cost distribution.

**Claim-level amounts, not aggregates.** Four of the six candidates report losses
aggregated over a policy period or a year. We need individual claim amounts to fit a
severity distribution and to apply a per-claim deductible.

### Why not freMTPL2

It was the initial suggestion and it is wrong for this question. It is the best-known
teaching dataset in this area, which is exactly why it was suggested, but familiarity
is not fitness. Third-party liability severities include bodily injury, and the public
files carry no deductible or limit information at all. Qiwei's inventory caught this;
we are following the evidence rather than the default.

### Loading

```{r}
#| label: load-data
install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/R/", type = "source")

library(CASdatasets)
data(pricingame)

pol   <- pg16trainpol    # 87,228 policies, exposure and risk characteristics
claim <- pg16trainclaim  # 4,568 claims, individual amounts

str(pol)
str(claim)
```

Two practical notes for whoever handles the data. Most categorical levels are
anonymised for confidentiality, so they carry no interpretable meaning and should not
be used to tell a story about driver or vehicle type. And the join between the two
tables is on the policy identifier, so policies with no claim will not appear in the
claim table at all.

---

## The ground-up gap {#sec-gu}

Status: **fixed as a limitation**. This is the most important assumption in the
project and it belongs on a slide.

Qiwei's matrix has a column labelled GU: verified ground-up repair cost, before any
deductible, coinsurance or limit is applied.

**That column is zero for all six candidates.** No public dataset in the inventory
publishes a verified ground-up loss.

This is not a reason to keep searching. It is a structural property of insurance data.
An insurer only observes losses that were reported to it, and a policyholder has no
reason to report a loss smaller than their deductible. The data are therefore already
**left-truncated** at an unknown threshold, and any distribution fitted to them is a
conditional distribution dressed up as an unconditional one.

The consequence for us is specific and quantifiable:

- Fitting a severity distribution to observed claims and treating it as ground-up
  **understates the true frequency** of small losses.
- That in turn **overstates the cost** of the low-deductible design, because the small
  losses that a low deductible would newly pick up are missing from the data.
- The bias therefore runs in a known direction, and it favours the high-deductible
  design.

Stating the direction of the bias is much stronger than stating that a bias exists.
It also answers one of the questions every student is told to be ready for: what
information is missing, censored, or aggregated.

The Track 4 brief lists discussion of censoring and truncation as required evidence.
So the gap Qiwei found is not an obstacle to the project. It is one of the deliverables.

---

## Model specification {#sec-model}

Status: **fixed**, subject to the diagnostics below coming out as expected.

We use a collective risk model. For a portfolio over one year,

$$S = \sum_{i=1}^{N} X_i$$

where $N$ is the number of claims, $X_i$ are individual ground-up loss amounts, and
$N$ and $X_i$ are assumed independent.

### Frequency

**Negative binomial**, with Poisson fitted alongside as a diagnostic.

The rationale is overdispersion. A Poisson model forces the variance to equal the
mean, and motor claim counts almost always show a variance above the mean because
policyholders are not identical. The negative binomial arises naturally as a Poisson
mixture over unobserved heterogeneity, which is the right story here given that the
risk characteristics in this dataset are anonymised and cannot fully explain that
heterogeneity.

Fit both, report the dispersion statistic, and keep Poisson in the appendix. If the
data turn out not to be overdispersed, we use Poisson and say so.

Exposure enters as an offset, since policies are observed for varying fractions of a
year.

### Severity

**Lognormal and gamma, choose by fit.**

Both are standard for positive, right-skewed loss amounts. Lognormal has the heavier
right tail of the two, which matters for a question about tail risk. Select by AIC
supported by a QQ plot, and show the QQ plot rather than only the AIC number, because
a number is not interpretable and a plot is.

We do **not** use a Pareto-type severity. It would be defensible, but with 4,568
claims the tail index would be poorly estimated, and the brief is explicit that a
simple analysis that is correctly understood beats a complex one that cannot be
defended.

### What we are not doing

No GLM with covariates. No machine learning. No traffic simulation. No correlated or
common-shock structure between claims. Each of these was considered and dropped for
scope; if asked, the answer is that the brief limits us to one main dataset and two
principal methods, and that the question does not require them.

```{r}
#| label: fit
library(fitdistrplus)
library(MASS)

# Frequency
fit_pois <- glm(ClaimNb ~ 1, offset = log(Exposure), family = poisson,       data = pol)
fit_nb   <- glm.nb(ClaimNb ~ 1 + offset(log(Exposure)),                      data = pol)

# Severity
fit_ln <- fitdist(claim$ClaimAmount, "lnorm")
fit_ga <- fitdist(claim$ClaimAmount, "gamma")

gofstat(list(fit_ln, fit_ga))
```

---

## Scenario multipliers {#sec-scenario}

Status: **fixed for the baseline, sensitivity grid open**.

We do not invent the frequency and severity shift. It comes from published insurance
loss data, which is what makes this project defensible rather than a parameter
guessing exercise.

The Highway Loss Data Institute compared 2017-22 model year vehicles with and without
a bundle of driver assistance features, holding the vehicle model otherwise constant.
Matt Moore of IIHS-HLDI published the comparison on 11 August 2026. The findings:

| Quantity | Effect for equipped vehicles |
|---|---|
| Collision claim severity | about **10% higher** |
| Collision claim frequency | about **10% lower** |
| Property damage liability claim frequency | about **40% lower** |
| Overall collision losses | about **5% lower** |

The mechanism is that sensors are expensive to replace and usually require
calibration afterwards, but equipped vehicles are in fewer crashes to begin with.

### Baseline scenario

Our data are third-party material damage, so the property damage liability figures are
the closer match.

- Frequency multiplier $\phi_N = 0.60$
- Severity multiplier $\phi_X = 1.10$

Applied as: $\lambda_{AV} = \phi_N \lambda$, and severity scaled by $\phi_X$ on the
mean while holding the coefficient of variation constant.

### Sensitivity grid

Run the full model over $\phi_N \in [0.5, 1.0]$ and $\phi_X \in [1.0, 1.3]$ and
present a contour of the result. This is the single most valuable slide in the deck,
because it converts our biggest weakness, dependence on assumed parameters, into the
subject of the analysis.

### Known weakness in this mapping

**Open**, and someone should own writing two sentences on it.

Property damage liability pays for damage to the *other* party's vehicle. Driver
assistance on the insured vehicle reduces the frequency of causing such damage, but
the severity of the repair applies to a third-party vehicle that may not be equipped.
Applying a severity uplift derived from collision coverage to a liability portfolio is
therefore an approximation. It is a reasonable one if the vehicle fleet as a whole is
becoming equipped, which is the scenario we are describing, but it should be stated
rather than hidden.

---

## Contract designs {#sec-contracts}

Status: **fixed in structure, parameters open pending calibration**.

Insurer payment on a ground-up loss $X$:

$$Y = \min\big(\alpha (X - d)_+,\; L\big)$$

where $d$ is the deductible, $\alpha$ the insurer's share above the deductible, and
$L$ the per-claim limit.

| | Design A: high deductible | Design B: coinsurance |
|---|---|---|
| Deductible $d$ | 1,000 | 250 |
| Insurer share $\alpha$ | 1.00 | 0.80 |
| Per-claim limit $L$ | same | same |

### The calibration that makes the comparison fair

**Calibrate the two designs to approximately equal expected insurer cost at baseline,
then compare their tail behaviour and their effect on the policyholder.**

Without this step the comparison is trivial and the presentation is weak: of course a
higher deductible costs the insurer less. Holding expected cost equal turns the
question into something that actually has an answer worth defending, which is how the
two designs distribute risk rather than how much they cost on average.

Tune $d$ in Design A, or $\alpha$ in Design B, until expected aggregate cost matches
within a small tolerance. Report the calibrated values and the tolerance achieved.

```{r}
#| label: payment
insurer_payment <- function(x, d, alpha, limit) {
  pmin(alpha * pmax(x - d, 0), limit)
}

simulate_aggregate <- function(n_sims, lambda, size, sev_par, d, alpha, limit) {
  vapply(seq_len(n_sims), function(i) {
    n <- rnbinom(1, mu = lambda, size = size)
    if (n == 0) return(0)
    x <- rlnorm(n, sev_par$meanlog, sev_par$sdlog)
    sum(insurer_payment(x, d, alpha, limit))
  }, numeric(1))
}
```

Use at least 100,000 simulations, set a seed, and report the Monte Carlo standard
error on the tail measures. A tail estimate quoted without its own uncertainty invites
an obvious question.

---

## Required outputs {#sec-outputs}

Status: **fixed**. Every item below appears either on a slide or in the appendix.

**Aggregate loss distribution**, both designs, both baseline and shifted scenarios:

- Expected aggregate insurer cost, $E[S]$
- Standard deviation
- VaR at 95% and 99%
- **TVaR at 99%** (the headline tail measure)

**Policyholder side:**

- Expected retained loss per policyholder, $E[X - Y]$
- Retention as a share of ground-up loss

**Truncation and censoring diagnostics** (required by the Track 4 brief):

- Share of claims falling entirely below the deductible under each design
- Share of claims hitting the limit
- How each design changes what the insurer would observe in future data

**Sensitivity:**

- Contour of the design preference over the $(\phi_N, \phi_X)$ grid

Keep every figure legible from the back of a room. Two well-labelled charts beat six
crowded ones, and the deck is capped at 20 slides including cover and figures.

---

## Section ownership {#sec-owners}

Five sections in the brief, five of us. Each person owns their section and presents
it, two to four minutes each, no more than four slides each excluding the cover.

| # | Section | Owner | Depends on |
|---|---|---|---|
| 1 | Problem and decision | **open** | nothing, can start now |
| 2 | Key concepts | Hoyong Jung | nothing, can start now |
| 3 | Analysis and evidence | Qiwei Ye | @sec-data, @sec-model |
| 4 | Comparison and judgment | Daren Wang | section 3 outputs |
| 5 | Conclusion and limitations | Xiaoxiao Fan | @sec-gu, @sec-scenario |

Section 1 is unclaimed. Faye, it is yours if you want it, and it is the section with
the fewest dependencies.

### A note for sections 4 and 5

Both are downstream of the simulation, which creates a scheduling risk: if you wait
for section 3 to finish, you compress into the final days.

Neither of you actually needs to wait.

**Section 4** can be written against the *structure* of the result rather than the
numbers. The comparison logic, which measures matter and why, and what would make each
design preferable is all fixed by @sec-contracts. Draft it with placeholder numbers.

**Section 5** can be drafted almost entirely from this document today. The limitations
register below is the raw material, and it will not change when the simulation
finishes.

### One open role

**Slide template owner: open.** The brief requires the template to be consistent
across all slides. Five people building slides independently produces five decks and
a bad evening on 8 September. Whoever takes this should push a template to the repo
before anyone starts making slides.

---

## Limitations register {#sec-limitations}

Raw material for section 5. Roughly in order of how badly each one would hurt if a
question landed on it.

1. **No ground-up losses.** No public dataset publishes verified ground-up repair
   cost. Observed claims are left-truncated at an unknown reporting threshold, so the
   fitted severity is conditional, small losses are undercounted, and the bias favours
   the high-deductible design. See @sec-gu.

2. **Coverage mismatch in the scenario.** Frequency and severity multipliers come from
   US collision and property damage liability experience; the data are French
   third-party material damage. Both the country and the coverage differ. See
   @sec-scenario.

3. **Driver assistance is not full automation.** HLDI measures vehicles with
   assistance bundles, not autonomous vehicles. We use it as the best available
   observed evidence for the direction and rough size of the shift, not as a
   measurement of automation.

4. **Independence assumed.** Claims are treated as independent. Vehicles sharing
   software or sensor hardware plausibly have correlated failure modes, which would
   fatten the aggregate tail beyond what we report. Our tail measures are therefore
   likely optimistic. We did not model this; it is the obvious next step.

5. **Anonymised categorical levels.** Risk characteristics in this dataset have no
   published meaning, so we cannot segment or interpret by risk class.

6. **Single year, single portfolio.** No trend, no inflation adjustment, no repair
   cost index. Repair costs have been rising faster than general inflation, which
   would push severity up further.

7. **No policyholder behaviour.** We compute retention mechanically. In reality a
   higher deductible changes reporting behaviour and lapse rates, which would feed
   back into both frequency and portfolio composition.

8. **Expenses and recoveries excluded.** No claim handling expense, salvage, or
   subrogation. All are absent from the public data, per Qiwei's inventory.

---

## Questions we will be asked {#sec-qa}

The brief lists the questions every student should be ready to answer. Roughly six
minutes of questions across the group, and the checklist requires each of us to
respond without relying on slide text. Everyone should be able to answer these for
their own section.

- What is the central question, and why is it worth investigating?
- Why a negative binomial rather than a Poisson?
- Why lognormal rather than gamma, or rather than Pareto?
- Where did the frequency and severity multipliers come from?
- What is the most important assumption, and which way does it bias the result?
- Why did you hold expected cost equal between the two designs?
- What does TVaR at 99% mean in plain language?
- What is missing from your data that you would want before using this in practice?

The second, third and sixth are the ones most likely to catch someone out, because
they are choices we made rather than facts we looked up.

---

## Timeline {#sec-timeline}

| Date | Milestone |
|---|---|
| Sun 31 Aug | Data loaded, distributions fitted, diagnostics agreed |
| Wed 3 Sept | Simulation running, baseline and scenario results produced |
| Fri 5 Sept | Analysis frozen. No new numbers after this point |
| Sun 7 Sept | Slides merged into one template, first full run-through |
| Tue 8 Sept | Final run-through with timing, question practice |
| Wed 9 Sept | Presentation, 18:00 EST |

The freeze on 5 September is the important one. Slides built on numbers that are still
moving get rebuilt, and there is no room to rebuild them.

---

## References

Sources to carry onto the final reference slide.

- Dutang, C. and Charpentier, A. *CASdatasets*: Insurance datasets.
  `pricingame` reference, French Institute of Actuaries Pricing Games.
  <https://dutangc.github.io/CASdatasets/reference/pricingame.html>
- Highway Loss Data Institute. Comparison of 2017-22 model year vehicles with and
  without driver assistance bundles. Analysis by M. Moore, IIHS-HLDI, 11 August 2026.
- Highway Loss Data Institute. *Compendium of HLDI collision avoidance research*,
  Bulletin Vol. 40, No. 4, April 2023.
- Ye, Q. *Public Insurance Dataset Candidates: Auto Deductible vs Coinsurance
  Simulation.* Six-dataset inventory with variable availability matrix, 21 August 2026.
- MFI Refresher Group Presentation Project, Student Project Brief, Track 4.
